---

---
# Reshaping {-} 


```{r}
library(tidyverse)
library(tidyr)
library(ggplot2)
library(dplyr)
```

```{r}
sleep_wide <- read.csv("https://mac-stat.github.io/data/sleep_wide.csv")

head(sleep_wide)
```

# units of observation are the subjects.

# I would need to create a column that holds all the reaction times and then a day column

#ggplot(sleep_wide, aes(y = rxn time, x = day, color = Subject))

```{r}
sleep_day_rxn <- sleep_wide |>
  pivot_longer(cols = -Subject, names_to = "day", names_prefix = "day_", values_to = "rxn_time")

ggplot(sleep_day_rxn, aes(y=rxn_time,x=day, color =Subject))+
  geom_line()
```

```{r}
sleep_long <- sleep_wide |>
  pivot_longer(cols = -Subject,
               names_to = "day",
               names_prefix = "day_",
               values_to = "reaction_time") |>
  mutate(Subject = as.factor(Subject),day = as.numeric(day))
head(sleep_long,3)
```

#changed also to make each of the 9 days of the subject study broken up

```{r}
ggplot(sleep_long, aes(y = reaction_time, x = day, color = Subject)) + 
  geom_line()
```

```{r}
# Make a line plot of reaction time by day for each subject
# Put these all on separate frames (one per subject)
ggplot(sleep_long, aes(y = reaction_time, x = day)) + 
  geom_line() + 
  facet_wrap(~ Subject)
```

# reaction time gets much worse around day 5 of lack of sleep for the majority of subjects, meaning less sleep results in a slower reaction time.

```{r}
# sleep_long |>
#   pivot_wider(names_from = , values_from = ___) |>
#   head()
```
